---
title: "R_Maps"
author: "JAT"
date: "28/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Natural earth R package

<https://www.naturalearthdata.com/>

```{r setup}

#devtools::install_github("ropenscilabs/rnaturalearthdata")
#install.packages("rnaturalearthhires",
#                 repos = "http://packages.ropensci.org",
#                type = "source")

#install.packages("rnaturalearth")

library("rnaturalearth")
library("rnaturalearthdata")


library(sp) #spatialPointsDataFrame

```


```{r}


vignette('what-is-a-country', package='rnaturalearth')
```


```{r all countries}


#world countries
sp::plot(ne_countries())
#sp::plot(ne_countries(continent="europe"))
```
```{r all countries}


#world countries
sp::plot(ne_states())
```

```{r one country}

#uk
sp::plot(ne_countries(country = 'united kingdom'))


# countries, UK undivided
#sp::plot(ne_countries(country = 'united kingdom', type = 'countries'))
```
```{r}
# map_units, UK divided into England, Scotland, Wales and Northern Ireland
sp::plot(ne_countries(country = 'united kingdom', type = 'map_units'))
```
```{r}
sp::plot(ne_states(country = 'spain')) 

```
```{r}
sp::plot(ne_states(country = 'turkey')) 

```


```{r regions}
#states, admin level1 boundaries
sp::plot(ne_states(country = 'spain')) 
```




```{r}
#lakes
#install.packages("rgdal")
lakes110 <- ne_download(scale = 110, type = 'lakes', category = 'physical')
sp::plot(lakes110)
```

## ggplot2 and naturalearth


```{r}
#install.packages(c("cowplot", "googleway", "ggplot2", "ggrepel","ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata"))
#install.packages("rgeos")

library("ggplot2")
theme_set(theme_bw())
library("sf")

library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

ggplot(data = world) +
    geom_sf() #it adds a geometry stored in a sf object
```

```{r}
ggplot(data = world) +
    geom_sf() +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("World map", subtitle = paste0("(", length(unique(world$name)), " countries)"))
```


```{r}
ggplot(data = world) + 
    geom_sf(color = "black", fill = "lightgreen")
```

```{r}
ggplot(data = world) +
    geom_sf(aes(fill = pop_est)) +
    scale_fill_viridis_c(option = "plasma", trans = "sqrt")
```

The function coord_sf allows to deal with the coordinate system, which includes both projection and extent of the map

```{r}

ggplot(data = world) +
    geom_sf() +
    coord_sf(crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs ")
```


The extent of the map can also be set in coord_sf, in practice allowing to “zoom” in the area of interest, provided by limits on the x-axis (xlim), and on the y-axis (ylim). Note that the limits are automatically expanded by a fraction to ensure that data and axes don’t overlap; it can also be turned off to exactly match the limits provided with expand = FALSE

```{r}
ggplot(data = world) +
    geom_sf() +
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)

```


```{r}
ggplot(data = world) +
    geom_sf() +
    coord_sf(xlim = c(-50, 50), ylim = c(25, 75), expand = FALSE)

```


```{r}
library("ggspatial")
ggplot(data = world) +
    geom_sf() +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = c(-50, 50), ylim = c(25, 75))

## Scale on map varies by more than 10%, scale bar may be inaccurate
```


```{r save map}

ggsave("map.pdf")
ggsave("map_web.png", width = 6, height = 6, dpi = "screen")

```


R World Map

```{r}

#install.packages("rworldmap")
library(rworldmap)

theCountries <- c("DEU", "COD", "BFA")
# These are the ISO3 names of the countries you'd like to plot in red

malDF <- data.frame(country = c("DEU", "COD", "BFA"),
  malaria = c(1, 1, 1))
# malDF is a data.frame with the ISO3 country names plus a variable to
# merge to the map data

malMap <- joinCountryData2Map(malDF, joinCode = "ISO3",
  nameJoinColumn = "country")
# This will join your malDF data.frame to the country map data

mapCountryData(malMap, nameColumnToPlot="malaria", catMethod = "categorical",
  missingCountryCol = gray(.8))
# And this will plot it, with the trick that the color palette's first
# color is red

```



```{r}
## Create multiple color codes, with Burkina Faso in its own group
malDF <- data.frame(country = c("DEU", "COD", "BFA"),
  malaria = c(1, 1, 2))

## Re-merge
malMap <- joinCountryData2Map(malDF, joinCode = "ISO3",
  nameJoinColumn = "country")

## Specify the colourPalette argument
mapCountryData(malMap, nameColumnToPlot="malaria", catMethod = "categorical",
  missingCountryCol = gray(.8), colourPalette = c("red", "blue"))

```




# SHAPEFILES

```{r}

#https://www.r-graph-gallery.com/168-load-a-shape-file-into-r.html

wd<-dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(wd)

# Download the shapefile.
download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" ,
destfile=paste(wd,"/SHAPEFILES/world_shape_file.zip",sep=""))

# Unzip this file.
zipF<- paste(wd,"/SHAPEFILES/world_shape_file.zip",sep="")
outDir<-paste(wd,"/SHAPEFILES",sep="")
unzip(zipF,exdir=outDir)

#  -- > You now have 4 files. One of these files is a .shp file! (TM_WORLD_BORDERS_SIMPL-0.3.shp)


# Read this shape file with the rgdal library. 
library(rgdal)
my_spdf <- readOGR( 
  dsn= paste0("./SHAPEFILES") , 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

# -- > Now you have a Spdf object (spatial polygon data frame). You can start doing maps!

# Basic plot of this shape file:
par(mar=c(0,0,0,0))
plot(my_spdf, col="#f2f2f2", bg="skyblue", lwd=0.25, border=0 )

```

```{r}
#with ggplot2

# 'fortify' the data to get a dataframe format required by ggplot2
library(broom)
spdf_fortified <- tidy(my_spdf, region = "NAME")

# Plot it
library(ggplot2)
ggplot() +
  geom_polygon(data = spdf_fortified, aes( x = long, y = lat, group = group), fill="#69b3a2", color="white") +
  theme_void() 


```

## maptools

```{r}
library(maptools)

data(wrld_simpl)
myCountries = wrld_simpl@data$NAME %in% c("Australia", "United Kingdom", "Germany", "United States", "Sweden", "Netherlands", "New Zealand")

plot(wrld_simpl, col = c(gray(.80), "red")[myCountries+1])
```



## mapSpain

<https://github.com/rOpenSpain/mapSpain>

```{r}
#install.packages("mapSpain", dependencies = TRUE)
library(mapSpain)

#https://datos.gob.es/es/catalogo/e00125901-siane-carto-base


census <- mapSpain::pobmun19

# Extract CCAA from base dataset

codelist <- mapSpain::esp_codelist

census <-
  unique(merge(census, codelist[, c("cpro", "codauto")], all.x = TRUE))

# Summarize by CCAA
census_ccaa <-
  aggregate(cbind(pob19, men, women) ~ codauto, data = census, sum)

census_ccaa$porc_women <- census_ccaa$women / census_ccaa$pob19
census_ccaa$porc_women_lab <-
  paste0(round(100 * census_ccaa$porc_women, 2), "%")

# Merge into spatial data

CCAA_sf <- esp_get_ccaa()
CCAA_sf <- merge(CCAA_sf, census_ccaa)
Can <- esp_get_can_box()


# Plot with ggplot
library(ggplot2)


ggplot(CCAA_sf) +
  geom_sf(aes(fill = porc_women),
    color = "grey70",
    lwd = .3
  ) +
  geom_sf(data = Can, color = "grey70") +
  geom_sf_label(aes(label = porc_women_lab),
    fill = "white", alpha = 0.5,
    size = 3,
    label.size = 0
  ) +
  scale_fill_gradientn(
    colors = hcl.colors(10, "Blues", rev = TRUE),
    n.breaks = 10,
    labels = function(x) {
      sprintf("%1.1f%%", 100 * x)
    },
    guide = guide_legend(title = "Porc. women")
  ) +
  theme_void() +
  theme(legend.position = c(0.1, 0.6))



```




## giscoR


```{r}


library(giscoR)

# Set the same resolution for a perfect fit

res <- "03"

# Same crs
target_crs <- 3035

all_countries <- gisco_get_countries(
  resolution = res,
  epsg = target_crs
)
eu_countries <- gisco_get_countries(
  resolution = res, region = "EU",
  epsg = target_crs
)
ccaa <- esp_get_ccaa(
  moveCAN = FALSE, resolution = res,
  epsg = target_crs
)

library(ggplot2)

ggplot(all_countries) +
  geom_sf(fill = "#DFDFDF", color = "#656565") +
  geom_sf(data = eu_countries, fill = "#FDFBEA", color = "#656565") +
  geom_sf(data = ccaa, fill = "#C12838", color = "grey80", lwd = .1) +
  # Center in Europe: EPSG 3035
  coord_sf(
    xlim = c(2377294, 7453440),
    ylim = c(1313597, 5628510)
  ) +
  theme(
    panel.background = element_blank(),
    panel.grid = element_line(
      colour = "#DFDFDF",
      linetype = "dotted"
    )
  )
```

```{r}


#http://blog.manugarri.com/making-a-beautiful-map-of-spain-in-ggplot2/


```




```{r}

## tmap

#<https://github.com/r-tmap/tmap>

#install.packages("tmap")
library(tmap)



```





```{r}


## GEOJSON

wd<-dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(wd)

#install.packages("geojsonR")
library(geojsonR)


file_js = FROM_GeoJson(url_file_string = "./feature_collection.geojson")

file_js

```


```{r}
<https://github.com/deldersveld/topojson>
  
```
